Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-05-15 paper2.qmd
section:5,6; process:write

Here, $s_{XI}$ represents the standard deviation for the individuals, $p_{IA}$ for the stimuli, $s_{ZJ}$ for the judges, and $p_{JK}$ for the judgments. We assume $s_{XI}$ varies depending on the teaching method group to which each student belongs. In the example from @sec-theoretical, where the teaching method $X_{I} = \{1,2\}$, the model sets the constraint according to Equation ([-@eq-mat56]). This constraint anchors the scale of the individuals' latent trait while relaxing the assumption of equal dispersion for the stimuli, thereby addressing the concerns raised in @sec-theory-issue1a. 
$$
\sum_{g=1}^{2} s_{XI}[g]/2 = 1
$$ {#eq-mat56}

In contrast, we define $p_{IA}$ as a proportion of $1$ to establish the scale of the stimuli' latent trait relative to the scale of the individuals' trait. Note that, since the error vector $\boldsymbol{e}$ follows an uncorrelated Multivariate Normal distribution, $e_{IA}$ follows a univariate Normal distribution by the properties of the distribution. As a result, $T_{IA}$ is also normally distributed, which effectively reverts to assuming Normal discriminal processes for the stimuli, an assumption originally proposed by @Thurstone_1927b (see @tbl-thurstone_cases)

@@

<!-- 6. Which are these new avenues for research? -->
Five research avenues deserve particular attention due to their direct impact on the reliability and validity of trait estimates and the accuracy of statistical inferences. First, researchers should investigate how sampling and comparison mechanisms influence CJ experimental outcomes and affect the robustness of statistical inferences. Second, further work is needed to understand how judges' biases impact the number of comparisons required to achieve reliable trait estimates. Third, researchers should re-evaluate existing methods for identifying and excluding 'misfitting' judges and stimuli to ensure their effectiveness. Fourth, studies should explore how the hierarchical structure of stimuli influences both experimental design and hypothesis testing in CJ contexts. Finally,  studies should explore principled approaches for selecting contextually relevant variables.

@@

<!-- 7.4 PPL and Bayesian Inference -->
Fourth and final, researchers must also be proficient in Probabilistic Programming Languages (PPLs) and Bayesian inference methods. This expertise is essential for translating context-specific conceptual models into bespoke statistical models that meaningfully address the research questions at hand. While such translation is not exclusive to Bayesian methods--frequentist approaches can also be used--we argue that Bayesian methods offer a more accessible pathway,  as they reduce the need for additional expertise in optimization theory and techniques.

>>>>

Here, $s_{XI}$ represents the standard deviation for the individuals, $p_{IA}$ for the stimuli, $s_{ZJ}$ for the judges, and $p_{JK}$ for the judgments. We assume $s_{XI}$ varies depending on the teaching method group to which each student belongs. Using the example from @sec-theoretical, where the teaching method $X_{I} = \{1,2\}$, the model sets the constraint according to Equation ([-@eq-mat56]). This constraint anchors the scale of the individuals' latent trait while relaxing the assumption of equal dispersion for the stimuli, thereby addressing the concerns raised in @sec-theory-issue1a. 
$$
\sum_{g=1}^{2} s_{XI}[g]/2 = 1
$$ {#eq-mat56}

Because the error vector $\boldsymbol{e}$ follows an uncorrelated Multivariate Normal distribution, the marginal distribution of $e_{IA}$ is a univariate Normal distribution with mean zero and standard deviation $p_{IA}$. We set $P_{IA}$ as a proportion of $1$ to establish the scale of the stimuli' latent trait relative to the scale of the individuals' trait. As a result, $T_{IA}$ is also normally distributed. This configuration effectively reinstates Thurstone’s original assumption of Normal discriminal processes for the stimuli (see @tbl-thurstone_cases).

@@

<!-- 6. Which are these new avenues for research? -->
Five research avenues merit particular attention for their direct impact on the reliability and validity of trait estimates and the accuracy of statistical inferences. First, researchers should investigate how sampling and comparison mechanisms influence CJ experimental outcomes and affect the robustness of statistical inferences. Second, further work is needed to understand how judges' biases impact the number of comparisons required to achieve reliable trait estimates. Third, researchers should re-evaluate existing methods for identifying and excluding 'misfitting' judges and stimuli to ensure their effectiveness. Fourth, studies should explore how the hierarchical structure of stimuli influences both experimental design and hypothesis testing in CJ contexts. Finally, researchers should explore principled approaches for selecting contextually relevant variables.

@@

<!-- 7.4 PPL and Bayesian Inference -->
Fourth and final, researchers must also be proficient in Probabilistic Programming Languages (PPLs) and Bayesian inference methods. This expertise is essential for translating context-specific conceptual models into bespoke statistical models that meaningfully address the research questions at hand. While such translations are not exclusive to Bayesian methods, we argue that Bayesian approaches offer a more accessible path. They reduce, for instance, the need for additional expertise in optimization techniques and theory often required by frequentist methods, thereby lowering the technical barrier for many applied researchers.
